Learning to Estimate Search Progress Using Graph Neural Networks \thanks{This work expands on \citet{sudry2022learning}.}

### 2024-11-05 report/submission.tex

These findings suggest that our GNN-based approach captures fundamental patterns in search behavior that are transferable across problem types. Additionally, our lightweight architecture achieved comparable or better performance than its more complex counterpart, indicating that the structural information captured by GNNs is more meaningful than architectural sophistication for this task.

@@

To make the dataset more practical for GNN training and to better simulate real-world scenarios, we implemented a two-stage pruning strategy. First, we filtered out graphs with more than 15,000 nodes to manage computational complexity. Second, we applied a dynamic pruning approach where for each remaining graph, we randomly selected a threshold $t \in \{0.3, 0.5, 0.7\}$ and kept only nodes with $\frac{\text{serial number}}{\text{total nodes}} < t$.

@@

Our implementation includes two distinct GNN architectures tailored for different application needs: a high-accuracy model optimized for performance and a lightweight version suited for resource-limited environments.

@@

\gur{GAT - global structure or only neighbors weighting?}

@@

\gur{Change - test, all data.}

@@

@@

\gur{Training details - train, eval, test.}

\gur{In the table report the error on the entire dataset, add a column of the test set error.}

The benchmark evaluations were conducted on a dataset of 11,751 filtered search trees, representing diverse problem instances across both domains. \gur{More details}

Table \ref{tab:performance} presents each approach's Mean Squared Error (MSE) on the full collection of nodes in the dataset. The traditional benchmarks showed varying performance levels, with VeSP achieving the worse score. The formula based benchmarks achieve an error magnitude of $10^{-1}$ down to $10^{-2}$, while the Random Forest model achieved the best performance with an MSE within the $10^{-2}$ range.

Our GNN-based approaches significantly outperformed all benchmarks. Both HeavyGNN and LightGNN achieved an MSE of $10^{-3}$, an order of magnitude improvement over the best-performing Random Forest model. Notably, despite its simplified architecture, LightGNN showed superior performance, suggesting that the core graph structural information is more important than complex architectural features for this task.

@@

\begin{enumerate}
    \item \textbf{Comparison with LSTM:} Our work did not directly compare the GNN-based approach with the LSTM-based method introduced by \citet{sudry2022learning}, and the results presented here could not be comparable to the dataset used in that work. Future research should conduct a direct comparison between these two approaches on a common dataset to provide a more comprehensive evaluation of their relative performance. \naomi{revise this}

>>>>

These findings suggest that our GNN-based approach captures fundamental patterns in search behavior that are transferable across problem types. Additionally, our lightweight architecture achieved comparable or better performance than its more complex counterpart, indicating that the structural information captured by GNNs is more meaningful than architectural sophistication for this task. \gur{update}

@@

To make the dataset more practical for GNN training and to better simulate real-world scenarios, we implemented a two-stage pruning strategy. First, we filtered out graphs with more than 15,000 nodes to manage computational complexity. Second, we applied a dynamic pruning approach where for each remaining graph, we randomly selected a threshold $t \in \{0.3, 0.5, 0.7\}$ and kept only nodes with $\frac{\text{serial number}}{\text{total nodes}} < t$. Due to the random nature of the pruning, the dataset slightly vary from run to run.

@@

We tested two distinct GNN architectures tailored for different application needs: a high-accuracy model optimized for performance and a lightweight version suited for resource-limited environments.

@@

@@

The data for the GNNs architectures was obtained from splitting the pruned dataset into training (70\%), validation (15\%), and test sets (15\%). The validation set was used for early stopping in the training process and the test set was used for model evaluation. The performance on each of the sets after training is reported in Table~\ref{tab:train-test}.

@@

The benchmark evaluations were conducted on a dataset of 11,751 filtered search trees, representing diverse problem instances across both domains.

Table~\ref{tab:performance} presents each approach's Mean Squared Error (MSE) on the full collection of nodes in the dataset. The traditional benchmarks showed varying performance levels, with VeSP achieving the worse score. The formula based benchmarks achieve an error magnitude of $10^{-1}$ down to $10^{-2}$, while the Random Forest model achieved the best performance with an MSE within the $10^{-2}$ range.

@@

Our GNN-based approaches significantly outperformed all benchmarks. Both HeavyGNN and LightGNN achieved an MSE of $10^{-3}$, an order of magnitude improvement over the best-performing Random Forest model. Notably, despite its simplified architecture, LightGNN showed superior performance, suggesting that the core graph structural information is more important than complex architectural features for this task. \gur{revise}

@@

\begin{enumerate}
    \item \textbf{Comparison with LSTM:} Our work did not directly compare the GNN-based approach with the LSTM-based method introduced by \citet{sudry2022learning}, and the results presented here could not be comparable to the dataset used in that work. Future research should conduct a direct comparison between these two approaches on a common dataset to provide a more comprehensive evaluation of their relative performance.
